---
title: "p8105_hw3_jc5929"
output: github_document
date: "2022-10-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse) 
library(ggridges) 
library(patchwork) 
library(dplyr)
library(knitr)
library(p8105.datasets)
```


**Question 1**

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)

```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))

instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```
Above is  a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()

```

A table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.


```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```

**Question 2**

```{r q2}

accel_data=read_csv("data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = 'activity_min',
    values_to = 'activity_count',
    names_prefix = 'activity_'
  )  %>%
mutate(activity_count=round(activity_count,2)) %>%
  mutate(day_type=
        case_when(
     day== 'Monday' ~ 'weekday',
     day== 'Tuesday' ~ 'weekday',
     day== 'Wednesday' ~ 'weekday',
     day== 'Thursday' ~ 'weekday',
     day== 'Friday' ~ 'weekday',
     day== 'Saturday' ~ 'weekend',
     day== 'Sunday' ~ 'weekend')
   ) %>%
  mutate(activity_min=as.numeric(activity_min),day=as.factor(day), week=
           as.integer(week))

```
The accel data set has `r nrow(accel_data)` rows and 6 columns. The data set 
contains 5 weeks of accelerometer data from a 63 year-old male with BMI 25. The
variables in this data set are week, day_id, day,activity_min, activity_count, 
and day_type. 

`day_id` is a `r class(pull(accel_data,day_id))` variable


`day` is a `r class(pull(accel_data,day))` variable


`week` is a `r class(pull(accel_data,week))` variable


`day_type` is a `r class(pull(accel_data,day_type))` variable


`activity_min` is a `r class(pull(accel_data,activity_min))` variable


`activity_count` is a `r class(pull(accel_data,activity_count))` variable

```{r}
accel_data %>%
  group_by(day_id,week,day) %>%
  summarize(n_obs=n(),
            total_activity=sum(activity_count)) %>%
  knitr::kable()
```

Activity counts appears to have the lowest on the two Saturdays (week4 and 
week5). 

```{r}
accel_data %>%
  mutate(activity_min=as.numeric(activity_min)) %>%
  ggplot(aes(x=activity_min,y=activity_count,color=day))+
  geom_line()+labs(title = "24-Hour Activity Count by Day",
    x = "Time",
    y = "Activity Count",
    caption = "Data from the accel data"
  ) + 
  scale_x_continuous(
    breaks = c(0, 360, 720, 1080, 1440), 
    labels = c("12AM", "6AM", "12PM", "6PM", "11:59PM"),
    limits = c(0, 1440)
    )

```

The plot shows that the activity count is lowest during sleeping hours, which is 
from 12am to 6am. And the individual appears to be more active in the afternoon,
from around 8pm to 10pm, especially on Fridays and Saturdays.

**Question 3**
```{r}
data("ny_noaa")
```
The ny_noaa data set contains `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` 
columns. Some key varibales inlcudes:

`id`: weather station id


`prcp`: Precipitation (tenths of mm)


`snow`: snowfall(mm)


`snwd`: snow depth (mm)


`tmax`: maximum temperature  (tenths of degrees C)


`tmin`: minimum temperature (tenths of degrees C)

All variables were converted into standard units. `prcp`, was measured in tenths
of mm, and was converted into mm.

The belows shows infromation of the missing values of the data set.
```{r}
 missing_noaa=ny_noaa %>%
  summarize(missing_prcp=sum(is.na(prcp)),
            missing_snow=sum(is.na(snow)),
            missing_snwd=sum(is.na(snwd)),
            missing_tmax=sum(is.na(tmax)),
            missing_tmin=sum(is.na(tmin))
            ) %>%
  knitr::kable()
missing_noaa

```
 From the table, it is observed that there is quite a lot of missing data for 
 `tamx` and `tmin`. tmax has about `r round(sum(is.na(ny_noaa$tmax))/nrow(ny_noaa),2)*100`% of 
 missing values, and missing values for tmin accounts approximately for 
 `r round(sum(is.na(ny_noaa$tmin))/nrow(ny_noaa),2)*100`%. `snow` and `snwd` has relatively 
 smaller proportion of missing values. If the data set has a large amount of 
 missing values, it will cause bias, and could influence the accuracy for analysis.
 
```{r}
ny_data= ny_noaa %>%
  janitor::clean_names() %>%
  mutate(date=as.character(date)) %>%
  separate(date,into=c('year','month','date'),sep='-') %>%
  mutate(tmax=as.numeric(tmax),
         tmin=as.numeric(tmin),
         year=as.factor(year),
         date=as.numeric(date),
         month=as.numeric(month), 
         prcp=prcp/10,
         tmax=tmax/10,
         tmin=tmin/10
         )

ny_data %>% 
  count(snow) %>%
  arrange(desc(n))

```
 
 All variables were converted into standard units. `prcp`, was measured in tenths
of mm, and was converted into mm. From the data set, it is seen that the most 
commonly observed value for snowfall is 0, which means that most days in nyc, it
is not snowing. Since snowy days only occur in some winter days.
 
```{r}
 
  ny_data %>%
  filter(month==01|month==07) %>%
  drop_na(tmax) %>%
  group_by(year,id,month) %>%
  summarize(tmax_average=mean(tmax)) %>%
  ggplot(aes(x=year,y=tmax_average,color=month)) + geom_point(alpha=0.3) +
  facet_grid(. ~ month)+labs(
    title = "Avg Max Temp in January and July in each NY Station Across Years",
    x = "Year",
    y = "Temp (°C)")+theme(axis.text.x = element_text(angle = 60, hjust = 1))


```
The average max temperature in January is within -10 to 10°C. The lowest turns 
out to be about -13.5°C. There is no significant outliers in January.

Average max temperature in July appears to be 25-30°C. There is one outlier where
the max temoerature is about 14°C. 

Comparing January and July, the average max temperature is lower than in July. 
It also appear to to fluctuate more in January, as the range in Junuary is more
taller than in July.

```{r}
plot1= ny_data %>%
    ggplot(aes(x=tmin,y=tmax))+geom_hex()+labs(
    title = "NY Temperatures (1981-2010)",
    x = "Min Temp (°C)",
    y = "Max Temp (°C)"
  )
  
plot2 = ny_data %>%
    filter(snow>0 & snow<100) %>%
    ggplot(aes(x=snow,fill=year))+geom_density(alpha=0.3,adjust=0.5,
                                               color='yellow')+ 
  labs(
    title = "NY Snowfall Values, 0-100mm (1981-2010)",
    x = "Snowfall (mm)",
  )
 
plot1+plot2
```
The left hand side plot shows `tmax` vs `tmin` for the data set. This plot shows
that most days have a max temperature of 25°C and a min temperature of 15°C.

The plot on the right shows the distribution of snowfall values greater than 0
and less than 100 separately by year. From the plot, we can see that there are
multiple peaks, thus it has a multimodal distrbition.  


