---
title: "p8105_hw3_jc5929"
output: github_document
date: "2022-10-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


**Question 1**

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)

```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))

instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```
Above is  a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()

```

A table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.


```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```

**Question 2**

```{r q2}
library(tidyverse) 
library(ggridges) 
library(patchwork) 
library(dplyr)
library(knitr)
accel_data=read_csv("data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = 'activity_min',
    values_to = 'activity_count',
    names_prefix = 'activity_'
  )  %>%
mutate(activity_count=round(activity_count,2)) %>%
  mutate(day_type=
        case_when(
     day== 'Monday' ~ 'weekday',
     day== 'Tuesday' ~ 'weekday',
     day== 'Wednesday' ~ 'weekday',
     day== 'Thursday' ~ 'weekday',
     day== 'Friday' ~ 'weekday',
     day== 'Saturday' ~ 'weekend',
     day== 'Sunday' ~ 'weekend')
   )
accel_data %>%
  group_by(day_id) %>%
  summarize(n_obs=n(),
            total_activity=sum(activity_count)) %>%
  knitr::kable()

  accel_data %>%
  group_by(week) %>%
  ggplot(aes(x=activity_min,y=activity_count,color=day_id))+
  geom_line()+geom_point(alpha=0.3)+viridis::scale_color_viridis(name='day_id')
```

**Question 3**
```{r}
library(p8105.datasets) 
data("ny_noaa")
ny_data= ny_noaa %>%
  janitor::clean_names() %>%
  mutate(date=as.character(date)) %>%
  separate(date,into=c('year','month','date'),sep='-') %>%
  mutate(tmax=as.numeric(tmax),tmin=as.numeric(tmin))

month_plot= 
    ny_data %>%
  filter(month=='01'|month=='07') %>%
  drop_na(tmax) %>%
  group_by(year,id,month) %>%
  summarize(tmax_average=mean(tmax)) %>%
  ggplot(aes(x=year,y=tmax_average,color=month)) +geom_point()+
  facet_grid(.~month)

month_plot

plot1= ny_data %>%
    ggplot(aes(x=tmin,y=tmax))+geom_hex()
  
plot2 = ny_data %>%
    filter(snow>0 & snow<100) %>%
    ggplot(aes(x=snow,fill=year))+geom_density(alpha=0.3,adjust=0.5,
                                               color='yellow')
 
plot1+plot2




```